Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2017-03-27 ident_manu.Rnw
results done for calibration exrapolation

\documentclass[preprint]{elsarticle}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage{lineno}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage{paralist}
\usepackage{bm}
\usepackage{fixltx2e}
% \usepackage[inline]{showlabels}

@@

# starting rmse
# 72.13955

@@

Calibration of parameter values for each of the seven selection heuristics produced model output that varied in precision and computational effort to identify optimal values (\cref{tab:calib}).  The seven evaluated subsets were those for \ac{do} in each of the four categories of parameters in \cref{tab:heurist1}, including the single parameter for temperature.  The subsets in \cref{tab:heurist2,tab:heurist3} for the parameters selected independent of category and equally within each category were also evaluated.  As expected, the number of required iterations to minimize \ac{rmse} of observed \ac{do} and model output varied in proportion to the number of parameters in a subset.  For example, only \Sexpr{toeng(temp$its)[[1]]} iterations were required to optimize the single parameter selected for the temperature category, whereas \Sexpr{top$its} iterations were required for the \Sexpr{toeng(top$nprm)[[1]]} parameters selected independent of category.  A slight positive assocation between collinearity and number of iterations was also observed, such that computational effort  to identify optimal values for a parameter set inreased with lower identifiability (higher $\gamma$).  Of the seven paramter subsets, selection independent of category and equally within each category produced the lowest \ac{rmse} values (\Sexpr{scinot(top$rmse)}, \Sexpr{scinot(cattop$rmse)}, respectively), whereas the optimal values for paramters in the Optics category had the lowest reduction in \ac{rmse} (\Sexpr{scinot(optics$rmse)}).  Lower \ac{rmse} was associated with higher average sensitivity of a parameter subset.  Parameters selected only in the optics category had the highest \ac{rmse} and lowest average L1 (\Sexpr{scinot(optics$avee)}), whereas parameters selected independent of category had the lowest \ac{rmse} and highest average L1 (\Sexpr{scinot(top$avee)}, \cref{tab:calib}).

% starting rmse of model without calibration
\cref{fig:areachg}
% does overall area change correlate with degree of sensitivity?

@@

\subsection{Implications of sensitivity and identifiability analyses}

@@

strt <- cals[[1]]$mat[1,] %>% 
  .[length(.)] %>% 
  scinot

totab <- caldat %>% 
  mutate(
    coll = ifelse(is.na(coll), 1, coll),
    coll = sapply(coll, scinot),
    nprm = sapply(nprm, scinot),
    its = sapply(its, scinot),
    rmse = sapply(rmse, scinot), 
    avee = sapply(avee, scinot)
    ) %>% 
  arrange(heurist)

cap_val <- 'Results of model calibration to observed \\ac{do} for each of seven parameter subsets selected by sensitivity within each parameter category (first), independent of category (second), and equally within each category (third). The collinearity values ($\\gamma$) and parameters of each subset are from \\cref{tab:heurist1,tab:heurist2,tab:heurist3}.  Mean L1 is the average of sensitivity values (\\cref{l1}) for the subset.  $RMSE$ shows the final error comparing observed and model output after the specified number of iterations for optimization.'
foot.val <- '\\footnotesize *Starting \\ac{rmse} $= 72.14$'
colnms <- c('$\\gamma$', '$n$ parameters', 'mean L1', 'Iterations', '$RMSE$*')

@@

<<areachg, eval = T, fig.cap = 'Estimated hypoxic area and monthly averages (km$^2 \\times$ 1000) using the default parameter settings (a, b) and the differences using a 50\\% increase in each parameter (c, d).  The selected parameters were the top sensitive in each parameter category for \\ac{do} in \\cref{fig:sensalltile,tab:dosens}. Error bars are 95\\% confidence intervals for the monthly averages in (b) and the difference in monthly averages (d) from the perturbed and default results.  Areal extent of estimated hypoxia on August 29\\textsuperscript{th} using the defaulat parameters (grey) and estimated changes with parameter perturbations is shown in (e).', fig.height = 7, fig.width = 7, echo = F, cache = TRUE, out.width = "\\textwidth">>=
# ##
# # area time series
# areats <- read_excel('ignore/runcompare.xlsx', sheet = 'area')
# 
# ##
# # get all arrays of hypox area for each parameter
# shts <- excel_sheets('ignore/runcompare.xlsx') %>% 
#   grep('array', ., value = T)
# latlon <- read_excel('ignore/runcompare.xlsx', sheet = 'Latlon') %>% 
#   list(
#     lat = c(.[,1])[[1]], 
#     lon = as.numeric(names(.)[-1])
#     ) %>% 
#   .[c('lat', 'lon')]
# 
# ls <- vector('list', length = length(shts))
# names(ls) <- shts
# for(sht in shts){
#  
#   cat(sht, '\t')
#   
#   # read the sheet, interpolate for higher res
#   tmp <- read_excel('ignore/runcompare.xlsx', sheet = sht, na = '-9999', col_names = F) 
#   tmp <- tmp[1:176,1:320] %>% 
#     as.matrix
#   obs <- list(x = latlon$lat, y = latlon$lon, z = tmp)
#   newgrd <- expand.grid(
#       x = approx(latlon$lat, n = 4 * ncol(tmp))$y,
#       y = approx(latlon$lon, n = 4 * nrow(tmp))$y
#       )
#   int_val <- fields::interp.surface(
#     obj = obs,
#     loc = newgrd
#     )
#   tmp <- data.frame(newgrd, tmp = int_val) %>%
#     spread(x, tmp) %>%
#     dplyr::select(-y) %>%
#     as.matrix %>% 
#     .[, ncol(.):1] %>% 
#     t
#   
#   # subset hypoxic areas, make polygon
#   # get only the large areas
#   tmp[tmp >= 64] <- NA
#   tmp[!is.na(tmp)] <- 1
#   tmp <- raster(tmp, 
#     xmn = min(latlon$lon), 
#     xmx = max(latlon$lon),   
#     ymn = min(latlon$lat), 
#     ymx = max(latlon$lat)
#     )
#   tmp <- rasterToPolygons(tmp, dissolve = T) %>% 
#     disaggregate
#   areas <- transform(data.frame(tmp), area = gArea(tmp, byid = T))
#   filt <- with(areas, which(area > 0.12)) #quantile(area, 0.99)))
#   tmp <- tmp[filt, ] %>% 
#     fortify
#   
#   ls[[sht]] <- data.frame(parm = sht, tmp)
#    
# }
# 
# # combine output from each sheet
# arrs <- do.call('rbind', ls) %>% 
#   mutate(
#     parm = gsub('_array$', '', parm)
#     ) %>% 
#   # filter(lat > 28) %>% 
#   tibble::remove_rownames(.) %>% 
#   unite('grp', parm, group, remove = F)
# 
# # save
# save(areats, arrs, file = 'data/area.RData', compress = 'xz')

@@

p5 <- ggmap(map) + 
  geom_polypath(data = toplo5, aes(x = long, y = lat, group = grp), alpha = 0.5) +
  geom_path(data = toplo6, aes(x = long, y = lat, group = grp, colour = parm), alpha = 0.8, size = 0.8) + 
  scale_y_continuous(expand = c(0, 0)) + 
  scale_x_continuous(expand = c(0, 0)) +
  scale_colour_manual(values = cols, labels = labs) +
  coord_equal() + 
  mytheme %+replace%
  theme(legend.position = 'none') +
  ggtitle(label = NULL, subtitle = '(e) Estimated areal extents on August 29, default extent in grey') + 
  scalebar(location = 'bottomleft', y.min = 28.35, y.max = 28.8, x.min = -93.3, x.max = -92, 
    dist = 20, dd2km= TRUE, model='WGS84', st.dist=.1, st.size = 2, height = 0.08)

>>>>

\documentclass[preprint]{elsarticle}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage{lineno}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage{paralist}
\usepackage{bm}
\usepackage{fixltx2e}
\usepackage[inline]{showlabels}

@@

# starting rmse
strt <- cals[[1]]$mat[1,] %>% 
  .[length(.)] %>% 
  scinot(., digits = 3)
# 72.13955

@@

Calibration of parameter values for each of the seven selection heuristics produced model output that varied in precision and computational effort to identify optimal values (\cref{tab:calib}).  The seven evaluated subsets were those for \ac{do} in each of the four categories of parameters in \cref{tab:heurist1}, including the single parameter for temperature.  The subsets in \cref{tab:heurist2,tab:heurist3} for the parameters selected independent of parameter category and equally within each category were also evaluated.  As expected, the number of required iterations of the optimization algorithm to minimize \ac{rmse} of observed \ac{do} and model output varied in proportion to the number of parameters in a subset.  For example, only \Sexpr{toeng(temp$its)[[1]]} iterations were required to optimize the single parameter selected for the temperature category, whereas \Sexpr{top$its} iterations were required for the \Sexpr{toeng(top$nprm)[[1]]} parameters selected independent of category.  A slight positive assocation between collinearity ($\gamma$) and number of iterations was also observed, such that computational effort  to identify optimal values for a parameter set increased with lower identifiability (higher $\gamma$).  Of the seven parameter subsets, selection independent of category and equally within each category produced the largest reduction in \ac{rmse} values from the default output (starting \ac{rmse} \Sexpr{strt}, reduction to \Sexpr{scinot(top$rmse, digits = 3)}, \Sexpr{scinot(cattop$rmse, digits = 3)}, respectively), whereas the optimal values for parameters in the Optics category had the smallest reduction (\Sexpr{scinot(optics$rmse, digits = 3)}).  Lower \ac{rmse} was associated with higher average sensitivity of a parameter subset.  Parameters selected only in the optics category had the highest \ac{rmse} and lowest average L1 (\Sexpr{scinot(optics$avee)}, \Sexpr{toeng(optics$nprm)} parameters), whereas parameters selected independent of category had the lowest \ac{rmse} and highest average L1 (\Sexpr{scinot(top$avee)}, \Sexpr{toeng(top$nprm)} parameters, \cref{tab:calib}).

<<>>=
data(area)

# areats
maxdef <- areats[which.max(areats$Original), ]

# areas with  months
areatxt<- areats %>% 
  dplyr::select(dts = 1, everything()) %>% 
  mutate(
    dts = as.Date(dts), 
    mo = month(dts, label = T)
  )

# default august area mean +- se
deftxt <- filter(areatxt, mo %in% 'Aug') %>% 
  .$Original %>% 
  t.test
deftxt <- with(deftxt, c(estimate, diff(conf.int)/2)) %>% 
  round

# mean difference and se from default
levs <- c('umax', 'ZKa', 'astar490', 'KG1', 'Tref')
labs <- paste0(c('umax', 'ZKa', 'astar490', 'KG1', 'Tref(nospA+nospZ)'), '_1')
labs <- par_txt(labs, frm = 'tex')
modiffs <- gather(areatxt, 'var', 'val', -dts, -mo, -Original) %>% 
  group_by(mo, var) %>% 
  summarise(
    se = diff(t.test(val, Original)$conf.int)/2,
    est = mean(val) - mean(Original)
  ) %>% 
  ungroup %>% 
  mutate(
    var = gsub('\\+50%', '', var), 
    var = factor(var, levels = levs, labels = labs)
  )

# umax July area mean inrease +- se
umaxtxt <- filter(modiffs, mo %in% 'Jul') %>% 
  filter(grepl('umax', var)) %>% 
  select(est, se) %>% 
  round

# tref July area mean decrease +- se
treftxt <- filter(modiffs, mo %in% 'Jul') %>% 
  filter(grepl('Tref', var)) %>% 
  select(est, se) %>% 
  round

load(file = 'data/sens_ests_cat_all.RData')
senstop <- sens_ests_cat_all[['O2']] %>% 
  group_by(Category) %>% 
  filter(error %in% max(error)) %>% 
  ungroup %>% 
  arrange(error) %>% 
  mutate(
    Parameter = par_txt(Parameter), 
    error = sapply(error, scinot)
  )
@

The estimated areal extent of bottom-water hypoxia using default parameter values for the 3-dimensional model peaked on August, 28\textsuperscript{th} at \Sexpr{scinot(maxdef$Original)} km$^2$ (\cref{fig:areachg}a, August mean area \Sexpr{deftxt[1]} $\pm$ \Sexpr{deftxt[2]} km$^2$, \cref{fig:areachg}b). Increases in the top sensitive parameters for \ac{do} in each category (\cref{tab:dosens}) increased the estimated areal extent of hypoxia (Figs. \ref{fig:areachg}c, \ref{fig:areachg}e), with the exception of the parameter for optimum temperature of growth (\Sexpr{par_txt('Tref(nospA+nospZ)_1')}), which caused a decrease (mean difference from July default area \Sexpr{treftxt$est} $\pm$ \Sexpr{treftxt$se} km$^2$, \cref{fig:areachg}d).  The largest increase in area was caused by an increase in the maximum growth rate of phytoplankton (\Sexpr{par_txt('umax_1')}), with the greatest increase occurring in July (monthly mean increase of \Sexpr{umaxtxt$est} $\pm$ \Sexpr{umaxtxt$se} km$^2$, \cref{fig:areachg}d).  Overall, changes in hypoxia extent were largest for parameters with large sensitivity values (i.e., L1 = \Sexpr{senstop[4, 2]} for \Sexpr{senstop[4, 1]}, L1 = \Sexpr{senstop[5, 2]} for \Sexpr{senstop[5, 1]}).

@@

@@

totab <- caldat %>% 
  mutate(
    coll = ifelse(is.na(coll), 1, coll),
    coll = sapply(coll, scinot),
    nprm = sapply(nprm, scinot),
    its = sapply(its, scinot),
    rmse = sapply(rmse, scinot, digits = 3), 
    avee = sapply(avee, scinot)
    ) %>% 
  arrange(heurist)

cap_val <- 'Results of model calibration to observed \\ac{do} for each of seven parameter subsets selected by sensitivity within each parameter category (first), independent of category (second), and equally within each category (third). The collinearity values ($\\gamma$) and parameters of each subset are from \\cref{tab:heurist1,tab:heurist2,tab:heurist3}.  Mean L1 is the average of sensitivity values (\\cref{l1}) for the subset.  $RMSE$ shows the final error comparing observed and model output after the specified number of iterations for optimization.'
foot.val <- '\\footnotesize *Starting \\ac{rmse} $= 72.140$'
colnms <- c('$\\gamma$', '$n$ parameters', 'mean L1', 'Iterations', '$RMSE$*')

@@

<<areachg, eval = T, fig.cap = 'Estimated hypoxic area and monthly averages (km$^2 \\times$ 1000) using the default parameter settings (a, b) and the differences using a 50\\% increase in each parameter (c, d).  The selected parameters were the top sensitive in each parameter category for \\ac{do} in \\cref{fig:sensalltile,tab:dosens}. Error bars are 95\\% confidence intervals for the monthly averages in (b) and the difference in monthly averages (d) from the perturbed and default results.  Areal extent of estimated hypoxia on August 28\\textsuperscript{th} using the defaulat parameters (grey) and estimated changes with parameter perturbations is shown in (e).', fig.height = 7, fig.width = 7, echo = F, cache = TRUE, out.width = "\\textwidth">>=
# ##
# # area time series
# areats <- read_excel('ignore/runcompare.xlsx', sheet = 'area')
# 
# ##
# # get all arrays of hypox area for each parameter
# shts <- excel_sheets('ignore/runcompare.xlsx') %>% 
#   grep('array', ., value = T)
# latlon <- read_excel('ignore/runcompare.xlsx', sheet = 'Latlon') %>% 
#   list(
#     lat = c(.[,1])[[1]], 
#     lon = as.numeric(names(.)[-1])
#     ) %>% 
#   .[c('lat', 'lon')]
# 
# ls <- vector('list', length = length(shts))
# names(ls) <- shts
# for(sht in shts){
#  
#   cat(sht, '\t')
#   
#   # read the sheet, interpolate for higher res
#   tmp <- read_excel('ignore/runcompare.xlsx', sheet = sht, na = '-9999', col_names = F) 
#   tmp <- tmp[1:176,1:320] %>% 
#     as.matrix
#   obs <- list(x = latlon$lat, y = latlon$lon, z = tmp)
#   newgrd <- expand.grid(
#       x = approx(latlon$lat, n = 4 * ncol(tmp))$y,
#       y = approx(latlon$lon, n = 4 * nrow(tmp))$y
#       )
#   int_val <- fields::interp.surface(
#     obj = obs,
#     loc = newgrd
#     )
#   tmp <- data.frame(newgrd, tmp = int_val) %>%
#     spread(x, tmp) %>%
#     dplyr::select(-y) %>%
#     as.matrix %>% 
#     .[, ncol(.):1] %>% 
#     t
#   
#   # subset hypoxic areas, make polygon
#   # get only the large areas
#   tmp[tmp >= 64] <- NA
#   tmp[!is.na(tmp)] <- 1
#   tmp <- raster(tmp, 
#     xmn = min(latlon$lon), 
#     xmx = max(latlon$lon),   
#     ymn = min(latlon$lat), 
#     ymx = max(latlon$lat)
#     )
#   tmp <- rasterToPolygons(tmp, dissolve = T) %>% 
#     disaggregate
#   areas <- transform(data.frame(tmp), area = gArea(tmp, byid = T))
#   filt <- with(areas, which(area > 0.12)) #quantile(area, 0.99)))
#   tmp <- tmp[filt, ] %>% 
#     fortify
#   
#   ls[[sht]] <- data.frame(parm = sht, tmp)
#    
# }
# 
# # combine output from each sheet
# arrs <- do.call('rbind', ls) %>% 
#   mutate(
#     parm = gsub('_array$', '', parm)
#     ) %>% 
#   # filter(lat > 28) %>% 
#   tibble::remove_rownames(.) %>% 
#   unite('grp', parm, group, remove = F)
# 
# # save
# save(areats, arrs, file = 'data/area.RData', compress = 'xz')

@@

p5 <- ggmap(map) + 
  geom_polypath(data = toplo5, aes(x = long, y = lat, group = grp), alpha = 0.5) +
  geom_path(data = toplo6, aes(x = long, y = lat, group = grp, colour = parm), alpha = 0.8, size = 0.8) + 
  scale_y_continuous(expand = c(0, 0)) + 
  scale_x_continuous(expand = c(0, 0)) +
  scale_colour_manual(values = cols, labels = labs) +
  coord_equal() + 
  mytheme %+replace%
  theme(legend.position = 'none') +
  ggtitle(label = NULL, subtitle = '(e) Estimated areal extents on August 28, default extent in grey') + 
  scalebar(location = 'bottomleft', y.min = 28.35, y.max = 28.8, x.min = -93.3, x.max = -92, 
    dist = 20, dd2km= TRUE, model='WGS84', st.dist=.1, st.size = 2, height = 0.08)

### 2017-03-27 ident_manu.tex
results done for calibration exrapolation

\usepackage{alltt}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage{lineno}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage{paralist}
\usepackage{bm}
\usepackage{fixltx2e}
% \usepackage[inline]{showlabels}

@@

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Sun Mar 26 15:32:37 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/b41a6df79d21b187457db88f058313b9b430416a}{b41a6df79d21b187457db88f058313b9b430416a}}

@@

Calibration of parameter values for each of the seven selection heuristics produced model output that varied in precision and computational effort to identify optimal values (\cref{tab:calib}).  The seven evaluated subsets were those for \ac{do} in each of the four categories of parameters in \cref{tab:heurist1}, including the single parameter for temperature.  The subsets in \cref{tab:heurist2,tab:heurist3} for the parameters selected independent of category and equally within each category were also evaluated.  As expected, the number of required iterations to minimize \ac{rmse} of observed \ac{do} and model output varied in proportion to the number of parameters in a subset.  For example, only six iterations were required to optimize the single parameter selected for the temperature category, whereas 850 iterations were required for the eight parameters selected independent of category.  A slight positive assocation between collinearity and number of iterations was also observed, such that computational effort  to identify optimal values for a parameter set inreased with lower identifiability (higher $\gamma$).  Of the seven paramter subsets, selection independent of category and equally within each category produced the lowest \ac{rmse} values ($67.05$, $67.67$, respectively), whereas the optimal values for paramters in the Optics category had the lowest reduction in \ac{rmse} ($72.14$).  Lower \ac{rmse} was associated with higher average sensitivity of a parameter subset.  Parameters selected only in the optics category had the highest \ac{rmse} and lowest average L1 ($2.81\times 10^{-4}$), whereas parameters selected independent of category had the lowest \ac{rmse} and highest average L1 ($0.03$, \cref{tab:calib}).

% starting rmse of model without calibration
\cref{fig:areachg}
% does overall area change correlate with degree of sensitivity?

@@

% calibration table, do
%latex.default(totab[, -1], file = "", rowlabel = "Selection heuristic",     caption = cap_val, caption.loc = "top", size = "scriptsize",     label = "tab:calib", rowname = totab$heurist, rgroup = c("1",         "2", "3"), insert.bottom = foot.val, n.rgroup = c(5,         1, 1), colheads = colnms)%
\begin{table}[!tbp]
{\scriptsize
\caption{Results of model calibration to observed \ac{do} for each of seven parameter subsets selected by sensitivity within each parameter category (first), independent of category (second), and equally within each category (third). The collinearity values ($\gamma$) and parameters of each subset are from \cref{tab:heurist1,tab:heurist2,tab:heurist3}.  Mean L1 is the average of sensitivity values (\cref{l1}) for the subset.  $RMSE$ shows the final error comparing observed and model output after the specified number of iterations for optimization.\label{tab:calib}} 
\begin{center}
\begin{tabular}{llllll}
\hline\hline
\multicolumn{1}{l}{Selection heuristic}&\multicolumn{1}{c}{$\gamma$}&\multicolumn{1}{c}{$n$ parameters}&\multicolumn{1}{c}{mean L1}&\multicolumn{1}{c}{Iterations}&\multicolumn{1}{c}{$RMSE$*}\tabularnewline
\hline
{\bfseries 1}&&&&&\tabularnewline
~~Optics&$3.87$&$3$&$2.81\times 10^{-4}$&$14$&$72.14$\tabularnewline
~~Organic Matter&$8.61$&$5$&$3.18\times 10^{-3}$&$154$&$68.7$\tabularnewline
~~Phytoplankton&$3.97$&$7$&$0.02$&$270$&$68.72$\tabularnewline
~~Temperature&$1$&$1$&$0.02$&$6$&$71.76$\tabularnewline
~~Zooplankton&$5.2$&$4$&$0.03$&$99$&$71.22$\tabularnewline
\hline
{\bfseries 2}&&&&&\tabularnewline
~~Independent of category&$6.81$&$8$&$0.03$&$850$&$67.05$\tabularnewline
\hline
{\bfseries 3}&&&&&\tabularnewline
~~Equally within category&$5.22$&$7$&$0.03$&$690$&$67.67$\tabularnewline
\hline
\end{tabular}\end{center}}
\footnotesize *Starting \ac{rmse} $= 72.14$\end{table}

@@

\caption{Estimated hypoxic area and monthly averages (km$^2 \times$ 1000) using the default parameter settings (a, b) and the differences using a 50\% increase in each parameter (c, d).  The selected parameters were the top sensitive in each parameter category for \ac{do} in \cref{fig:sensalltile,tab:dosens}. Error bars are 95\% confidence intervals for the monthly averages in (b) and the difference in monthly averages (d) from the perturbed and default results.  Areal extent of estimated hypoxia on August 29\textsuperscript{th} using the defaulat parameters (grey) and estimated changes with parameter perturbations is shown in (e).}\label{fig:areachg}
\end{figure}

>>>>

\usepackage{alltt}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage{lineno}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage{paralist}
\usepackage{bm}
\usepackage{fixltx2e}
\usepackage[inline]{showlabels}

@@

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Sun Mar 26 16:14:21 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/ad5c4632bfb7084298ee0ce337870bbac1db811c}{ad5c4632bfb7084298ee0ce337870bbac1db811c}}

@@

Calibration of parameter values for each of the seven selection heuristics produced model output that varied in precision and computational effort to identify optimal values (\cref{tab:calib}).  The seven evaluated subsets were those for \ac{do} in each of the four categories of parameters in \cref{tab:heurist1}, including the single parameter for temperature.  The subsets in \cref{tab:heurist2,tab:heurist3} for the parameters selected independent of parameter category and equally within each category were also evaluated.  As expected, the number of required iterations of the optimization algorithm to minimize \ac{rmse} of observed \ac{do} and model output varied in proportion to the number of parameters in a subset.  For example, only six iterations were required to optimize the single parameter selected for the temperature category, whereas 850 iterations were required for the eight parameters selected independent of category.  A slight positive assocation between collinearity ($\gamma$) and number of iterations was also observed, such that computational effort  to identify optimal values for a parameter set increased with lower identifiability (higher $\gamma$).  Of the seven parameter subsets, selection independent of category and equally within each category produced the largest reduction in \ac{rmse} values from the default output (starting \ac{rmse} $72.14$, reduction to $67.052$, $67.665$, respectively), whereas the optimal values for parameters in the Optics category had the smallest reduction ($72.139$).  Lower \ac{rmse} was associated with higher average sensitivity of a parameter subset.  Parameters selected only in the optics category had the highest \ac{rmse} and lowest average L1 ($2.81\times 10^{-4}$, three parameters), whereas parameters selected independent of category had the lowest \ac{rmse} and highest average L1 ($0.03$, eight parameters, \cref{tab:calib}).

The estimated areal extent of bottom-water hypoxia using default parameter values for the 3-dimensional model peaked on August, 28\textsuperscript{th} at $1.19\times 10^{4}$ km$^2$ (\cref{fig:areachg}a, August mean area 8988 $\pm$ 551 km$^2$, \cref{fig:areachg}b). Increases in the top sensitive parameters for \ac{do} in each category (\cref{tab:dosens}) increased the estimated areal extent of hypoxia (Figs. \ref{fig:areachg}c, \ref{fig:areachg}e), with the exception of the parameter for optimum temperature of growth (\textit{Tref(nospA+nospZ)$_{p1}$}), which caused a decrease (mean difference from July default area -1070 $\pm$ 1161 km$^2$, \cref{fig:areachg}d).  The largest increase in area was caused by an increase in the maximum growth rate of phytoplankton (\textit{umax}), with the greatest increase occurring in July (monthly mean increase of 4109 $\pm$ 1085 km$^2$, \cref{fig:areachg}d).  Overall, changes in hypoxia extent were largest for parameters with large sensitivity values (i.e., L1 = $0.05$ for \textit{ZKa}, L1 = $0.05$ for \textit{umax}).

@@

% calibration table, do
%latex.default(totab[, -1], file = "", rowlabel = "Selection heuristic",     caption = cap_val, caption.loc = "top", size = "scriptsize",     label = "tab:calib", rowname = totab$heurist, rgroup = c("1",         "2", "3"), insert.bottom = foot.val, n.rgroup = c(5,         1, 1), colheads = colnms)%
\begin{table}[!tbp]
{\scriptsize
\caption{Results of model calibration to observed \ac{do} for each of seven parameter subsets selected by sensitivity within each parameter category (first), independent of category (second), and equally within each category (third). The collinearity values ($\gamma$) and parameters of each subset are from \cref{tab:heurist1,tab:heurist2,tab:heurist3}.  Mean L1 is the average of sensitivity values (\cref{l1}) for the subset.  $RMSE$ shows the final error comparing observed and model output after the specified number of iterations for optimization.\label{tab:calib}} 
\begin{center}
\begin{tabular}{llllll}
\hline\hline
\multicolumn{1}{l}{Selection heuristic}&\multicolumn{1}{c}{$\gamma$}&\multicolumn{1}{c}{$n$ parameters}&\multicolumn{1}{c}{mean L1}&\multicolumn{1}{c}{Iterations}&\multicolumn{1}{c}{$RMSE$*}\tabularnewline
\hline
{\bfseries 1}&&&&&\tabularnewline
~~Optics&$3.87$&$3$&$2.81\times 10^{-4}$&$14$&$72.139$\tabularnewline
~~Organic Matter&$8.61$&$5$&$3.18\times 10^{-3}$&$154$&$68.701$\tabularnewline
~~Phytoplankton&$3.97$&$7$&$0.02$&$270$&$68.722$\tabularnewline
~~Temperature&$1$&$1$&$0.02$&$6$&$71.759$\tabularnewline
~~Zooplankton&$5.2$&$4$&$0.03$&$99$&$71.223$\tabularnewline
\hline
{\bfseries 2}&&&&&\tabularnewline
~~Independent of category&$6.81$&$8$&$0.03$&$850$&$67.052$\tabularnewline
\hline
{\bfseries 3}&&&&&\tabularnewline
~~Equally within category&$5.22$&$7$&$0.03$&$690$&$67.665$\tabularnewline
\hline
\end{tabular}\end{center}}
\footnotesize *Starting \ac{rmse} $= 72.140$\end{table}

@@

\caption{Estimated hypoxic area and monthly averages (km$^2 \times$ 1000) using the default parameter settings (a, b) and the differences using a 50\% increase in each parameter (c, d).  The selected parameters were the top sensitive in each parameter category for \ac{do} in \cref{fig:sensalltile,tab:dosens}. Error bars are 95\% confidence intervals for the monthly averages in (b) and the difference in monthly averages (d) from the perturbed and default results.  Areal extent of estimated hypoxia on August 28\textsuperscript{th} using the defaulat parameters (grey) and estimated changes with parameter perturbations is shown in (e).}\label{fig:areachg}
\end{figure}
